---
title: Implementation of Decision Analysis to evaluate different agroforestry interventions
  in Northern Vietnam
author: "Hoa Do, Cory Whitney, Eike Luedeling"
date: "`r Sys.Date()`"
output:
  html_document: default
---

This document illustrates the development of mathematical model to calcualte Net Present Values (NPV) of different agroforestry interventions promoted in Northwest Vietnam under the framework of ICRAF's project. The calculation of NPV for the decision to implement coffee-based agroforesty system with the combination of teak, plum, coffee, annual crop and grass against conventional monoculture of maize will be explained as the illustrative example for the whole process. The same function can be used for the simulationn of other interventions with corresponding input data table for each system provided along in Supplementary materials.

We employed `decisionSupport` R package and its functions to develop the model and to perform probabilistic simulation. 

The `chance_event()` function calculated the time series occurrence of risk events. This function required two main arguments specified in input tables which are the probability that a particular risk event will happen in a given year and the time line for simulation. 

The `gompertz_yield()` calculated the yields of tree crops using identified parameters in the assumed pattern.

The value varier function `vv()` produces time series that contain variation from a specified mean and a desired coefficient of variation, which both were parameters in the input tables. 

The `discount()` function was used as the last step of the decision model to calculate NPV. The function implemented discounting of future costs or benefits by a decision maker-specific discount rate, which was specified in input tables. 

`decisionSupport()` function conducted the full decision analyses by performing probabilistic simulation. The function in requires two inputs:

+ a data table (.csv format) specifying the names and distributions for all uncertain variables

+ an R function that predicts decision outcomes based on the variables named in the data table

We called out `decisionSupport` package in R and defined sample values for all variables:

```{r}
library(decisionSupport)
make_variables<-function(est,n=1)
{x<-random(rho=est,n=n)
for(i in colnames(x))assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}
```

Applying `make_variables` function to the input data table to generate one random number for each variable.

```{r}
make_variables(estimate_read_csv(paste("Input tables/CAFS_teak.csv",sep="")))
```

The mathematical model could then be coded. We named the function calculating NPV `AF_benefit`

```{r}
AF_benefit<-function(x,varnames)
{
  #We defined each variables as vectors of 25 values corresponding to 25 years of simunlation 
  
  mono_costs<-rep(0,n_years)# the annual cost of maize monoculture 
  establisment_cost<-rep(0,n_years)# costs for setting up agroforestry system
  maintenance_cost<-rep(0,n_years)# the annual cost spent in agroforestry system
  timber_yield<-rep(0,n_years)# the yield of teak
  tree_yield<-rep(0,n_years)# the yield of plum
  coffee_yield<-rep(0,n_years)# the yield of coffee
  AF_maize<-rep(0,n_years)# the yield of maize in agroforestry system
  mono_maize<-rep(0,n_years)# the yield of maize in monoculture stand
  grass_benefit<-rep(0,n_years)# the benefit from grass
  annual_benefit<-rep(0,n_years)# the benefit of annual crops
  reduced_soil_loss<-rep(0,n_years)# amount of soil loss reduced in agroforestry system compare to monoculture system
  
  #Simulating the chance for different events to occur during simulation period
  frost<-chance_event(coffee_risk,value_if = 1,n=n_years)
  tree.risk<-chance_event(tree_risk,value_if = 1,n=n_years)
  maize.risk<-chance_event(maize_risk,value_if = 1,n=n_years)
  
  #Number of timber harvest throughout 25 years of simulation
  timb_harvest<-as.numeric(1:n_years%%round(freq_tim_harvest)==0)
  
  #calculate system cost
  maintenance_cost[1]<-0
  maintenance_cost[2:4]<-vv(main_cost_stage1,CV_cost,3)/cur_change # converting currency from VND to USD using currency exchange rate (cur_change) and was applied for any variable with monetary value (i.e. costs and benefits)
  maintenance_cost[5:10]<-vv(main_cost_stage2,CV_cost,6)/cur_change
  maintenance_cost[11:25]<-vv(main_cost_stage3,CV_cost,15)/cur_change
  
  establisment_cost[1]<-establis_cost/cur_change
  establisment_cost[2:25]<-0


  #calculate system benefit
  
  #benefit of timber
  timber_benefit<-(timb_benefit*timb_harvest)/cur_change

  
  #benefit of tree
  tree_yield<-gompertz_yield(max_harvest = max_tree_harvest,
                             time_to_first_yield_estimate = time_first_tree_est,
                             time_to_second_yield_estimate = time_sec_tree_est,
                             first_yield_estimate_percent = first_tree_est_per,
                             second_yield_estimate_percent = sec_tree_est_per,
                             n_years=n_years,
                             var_CV=CV_tree_yield,
                             no_yield_before_first_estimate = FALSE)
  tree_yield<-tree_yield*(1-tree.risk*yield_tree_risk)
  tot_tree_yield<-tree_yield*num_of_tree
  tree_benefit<-(tot_tree_yield*tree_price)/cur_change
  

  
  #benefit of coffee
  coffee_yield[1:15]<-gompertz_yield(max_harvest = max_coff_harvest,
                                     time_to_first_yield_estimate = time_first_coff_est,
                                     time_to_second_yield_estimate = time_sec_coff_est,
                                     first_yield_estimate_percent = first_coff_est_per,
                                     second_yield_estimate_percent = sec_coff_est_per,
                                     n_years=15,
                                     var_CV = CV_coff_yield,
                                     no_yield_before_first_estimate = FALSE)
  coffee_yield[16:25]<-coffee_yield[6:15]
  coffee_yield<-coffee_yield*(1-frost*coff_frost_yield) # account risk of frost in the calculation of coffee yield
  tot_coffee_yield<-coffee_yield*num_coff_tree
  coffee_benefit<-(tot_coffee_yield*vv(coff_price,CV_coff_price,n_years))/cur_change

  
  #benefit of maize in AF system
 time<-1:n_years
  decay_speed_AF<--log(1-decay_rate_AF)
  AF_maize<-min_AFmaize+(max_AFmaize-min_AFmaize)*exp(-decay_speed_AF*(time-1))
  tot_AF_maize<-vv(AF_maize*(1-maize.risk*yield_maize_risk),CV_maize_yield,n_years)
  AF_maize_benefit<-(tot_AF_maize*vv(maize_price,CV_maize_price,n_years))/cur_change

  
  #benefit of miaze in monoculture
  decay_speed_mono<--log(1-decay_rate_mono)
  mono_maize<-max_monomaize*exp(-decay_speed_mono*(time-1))
  tot_mono_maize<-vv(mono_maize*(1-maize.risk*yield_maize_risk),CV_maize_yield,n_years)
  mono_revenue<-(tot_mono_maize*vv(maize_price,CV_maize_price,n_years))/cur_change
  mono_costs<-vv(mono_cost,CV_cost,n_years)/cur_change
  mono_benefit<-mono_revenue-mono_costs

  
  #benefit of grass
  grass_benefit<-vv(grass_profit,CV_grass_profit,25)/cur_change
  
  #benefit of soybean
  annual_benefit[1:3]<-annual_benefit/cur_change
  annual_benefit[4:25]<-0
  
  #benefit of soil erosion control
  reduced_soil_loss[1]<-0
  reduced_soil_loss[2:10]<-reduced_percents1*mono_soil_loss
  reduced_soil_loss[10:25]<-reduced_percents2*mono_soil_loss
  soil_saved<-vv(reduced_soil_loss,CV_eros_control,n=n_years)
  erosion_benefit<-(soil_saved*soil_saved_payment)/cur_change

  #Calculating NPVs of agroforestry and monoculture and NPV of the decision choosing agroforestry over monoculture of maize
system_benefit<-timber_benefit+tree_benefit+coffee_benefit+AF_maize_benefit+grass_benefit+annual_benefit+erosion_benefit
  bottomline_benefit<-system_benefit-maintenance_cost-establisment_cost
  tradeoff_benefit<-bottomline_benefit-mono_benefit
  cash_flow<-discount(bottomline_benefit,discount_rate=discount_rate,calculate_NPV = FALSE)# this returned 25 values of discounted cashflows for 25 years
  cum_cash_flow<-cumsum(cash_flow)# this calculated cummulative values of discounted cashflows
  NPV_mono<-discount(mono_benefit,discount_rate=discount_rate,calculate_NPV = TRUE )#NPV of monoculture system
  NPV_tradeoff<-discount(tradeoff_benefit,discount_rate=discount_rate,calculate_NPV = TRUE )#NVP of the decision choosing agroforestry over monoculture
  NPV_system<-discount(bottomline_benefit,discount_rate=discount_rate,calculate_NPV = TRUE)#NVP of agroforestry system
  return(list(cashflow=cum_cash_flow,trade_off=NPV_tradeoff,NPV_maize=NPV_mono,NPV=NPV_system))
}

```

Probabilistic simulation was implemented through `decisionSupport` function using input data table and the developed R function. This step returns Monte Carlo dataset with 10,000 values for each of input variables and each of response variables that are called out in the return list and VIP score for each input variable from PLS regression.

```{r message=FALSE,warning=FALSE}
decisionSupport(inputFilePath = paste("Input tables/CAFS_teak.csv",sep=""),
                outputPath = paste("MCResults/CAFS_teak",sep=""),
                write_table = TRUE,
                welfareFunction=AF_benefit,
                numberOfModelRuns=10000,
                functionSyntax="plainNames")

```





